Quasars Probing Quasars IX. The Kinematics of the Circumgalactic Medium Surrounding $z\sim2$ Quasars

### 2018-01-26 QPQ9/Submit/response_2.txt

Dear referee,

We thank you very much for careful reading of our manuscript the second time. Your constructive criticism has been
very helpful.

>=== Major concern 1: failure to capture velocity spread of the absorption and misuse of the central limit theorem
>
>In response to one of my concerns, the authors now motivate Gaussian fitting to the mean absorption profile using the
>central limit theorem. This is a fundamental misuse of the central limit theorem. The central limit theorem implies
>(under fairly loose assumptions) that the sums of random variables will tend towards a normal distribution. This means,
>for example, that the measured flux in a particular pixel of the stack will be normally distributed. But it *does not*
>imply that the mean line profiles themsleves will be Gaussian. While there are conditions under which mean line
>profiles will converge to a Gaussian, these are significantly more demanding conditions that typically do not apply to
>astrophysical measurements and objects. For example, the mean stack of point sources in ground based images will not
>converge to a Gaussian because of the large fraction of flux contained in wings of the PSF. Similarly, mean line
>profiles in stacks are often far from Gaussian such as: both broad and narrow quasar emission lines (Schneider+2010,
>Shen+2016b), galaxy emission lines (Cicone+2016), and associated quasar absorption lines which are distinctly
>non-Gaussian as shown in Figure 9 of this paper.
>
>Previous QPQ papers (Prochaska & Hennawi 2009, Lau+2016) and those from other authors (e.g. Tripp+2011, Johnson+2015,
>Rudie+2017) have shown that AGN and quasar hosts often exhibit absorbing gas at "extreme" velocities of 500-1000 km/s.
>Indeed, these extreme velocities are discussed in the introduction. However, it is clear that the extreme velocities
>are not being captured by the measured Gaussian fit.
>
>This concern is raised not only by the previous works but also by careful examination of Figure 2. In particular the
>pseudo-continuum level in the mean stack is systematically lower than in the median stack for all three ions
>considered. This indicates that there is additional absorption at large velocities which is not being captured by the
>analysis in the paper but is instead being erroneously folded into the pseudo-continuum estimate.
>
>This is a major issue with the adopted methodology that must be addressed with careful thought, new
>experiments/testing, and detailed discussion.

We thank the referee for pointing out the misuse of central limit theorem. It applies to the fluxes at each pixel.
The absorption in the stack will follow the distribution from which it was sampled, by the law of large numbers. We
present a more careful argument for fitting one Gaussian component and one constant component, and removed the tex
relying on the central limit theorem.

The distribution of absorbers unassociated with the foreground quasars is expected to be uniform and random, and
depress the continuum level in the stack uniformly. This justifies our fitting of a pseudocontinuum, and is not
the quantity of interest. The questions to be addressed is, whether one component is a good description of the
dispersion in the average absorption, and if so, whether the dispersion in this component is well captured by a
Gaussian. We tried replacing a constant pseudocontinuum with a second Gaussian absorption component in our modeling
of the CII stack. We found this second Gaussian component to be weak and broad. The amplitude is 0.03, compared to our
original constant component of 0.97. The standard deviation is 3600 km/s, whose span is wider than the entire
velocity window for stacking. This weak and very broad component also has little effect on the centroid and width of
the narrower Gaussian component. We therefore see no convincing evidence for a broader component hidden in the
depressed continuum. We then calculated the dispersion in a +/-1300 km/s window surrounding the absorption centroid
using the standard deviation formula. We hae this large velocity window to encompass most of the associated absorption.
We found the dispersion measured this way is essentially the same as that found by fitting a Gaussian. This also leans
support to our approach of fitting two Gaussians to the CIV and MgII doublet, where we cannot directly apply the
standard deviation formula.

In our revised manuscript, we acknowledge the possibility our Gaussian model does not capture all the powers at extreme
velocities. However we also note that occasional large kinematic offsets do not make a strong case against
gravitational motions. Firstly, quasars occasionally inhabit extremely overdense environments, such as
protoclusters. Secondly, the probability of intercepting a random absorber is boosted by large scale clustering. Hence
occasional extreme velocities would not necessitate outflows. Our preferred kinematic measure is thus the dispersion in
the average absorption. We added the above text to the revised manuscript.

The uniformly depressed continuum results from absorbers unrelated to the foreground quasars, which is expected to
have a covering fraction less than 50%. For median stack, a pixel is not affected unless there are more than 50% of
quasar pairs with absorption at its velocity. The pseudo-continuum of the mean stack being lower than the median stack
is not a reflection of associated absorption at large velocities.

>=== Major concern 2: insufficient detail in how the doublet fitting is performed.
>
>I thank the authors for specifying the allowed range of doublet ratio. However, additional information is still needed
>on the doublet fitting, particularly regarding the pseudo continuum.
>
>The doublet ratio is degenerate with centroid in some cases, particularly for CIV and this possible degeneracy is not
>discussed.
>
>Depending on how the pseudo continuum fitting is performed, it could potentially suppress any broad wings in the mean
>absorption profile. The authors have most likely taken steps to prevent this from happening. But if so, these steps are
>note specified.f
>
>Provide additional details addressing both of these points.

We tried fitting the continuum separately in (-3000,-1300) km/s and (+1300, +3000) km/s instead of multiplying a
constant to a Gaussian. We found the continuum determined this way is essentially the same.

We acknowledge the degeneracy between the centroid and the doublet ratio for CIV. And hence, we state that the
results for the doublets are presented for consistency check, but the analysis focuses on CII. We emphasized this
point in the revised manuscript. For CIV, a doublet ratio
closer 2:1 is preferred. We found a doublet ratio of 2:1 captures the absorption at the central few pixels better than
a doublet ratio of 1:1 does.

>=== Main concern 3: mis-representation of papers in the literature
>
>While most of these have been corrected, one remains and a new one has been added. In particular, both Tripp+2011 and
>Rudie+2017 found absorbing gas spread over ~1000 km/s which is comparable to the velocity spreads from the QPQ series
>that are discussed as extreme in the introduction. The main conclusions of Tripp+2011 and Rudie+2017 focus on these
>large velocity spreads and the current description in the introduction is therefore misleading.

We thank the referee for scrutinizing our representation of the literature. We dropped the words that claim these
absorption systems have most of the gas within several hundred km/s.

>=== In addition:
>
>A) To more convincingly demonstrate that the velocity shift in the measured mean absorption profile is real the authors
>should:
>
>1) Remove possible contamination from absorption systems associated with the background quasar. The paper stacks the
>background spectra for pairs in which the redshift difference between the foreground and background quasar is
>>3000 km/s in order to prevent contamination by physical quasar pairs in the presence of redshift uncertainty. However,
>quasars also exhibit narrow associated absorption systems at large velocities of up to ~12,000 km/s (Wild+2008) which
>could also artificially introduce an asymmetry into the experiment. The authors can easily demonstrate that this is not
>the case by repeating the stacking procedure on quasars pairs with a more restrictive range of allowed velocity
>separation of >12,000 km/s.
>
>2) Demonstrating that CII* cannot bias the measurement of the CII velocity centroid. Previous QPQ papers have
>identified CII* absorption in the CGM of quasars. This absorption feature is blue-shifted from the ground state
>transition by 260 km/s which is comparable to the measured offset. With data from Lau+2016, the authors can easily
>demonstrate that contamination from CII* is not responsible for the observed shift.

We identified and removed all BAL and mini-BAL systems in the background quasars. Possible contamination from narrow
associated absorbers of background quasars should result in a tilt in the pseudo-continuum. Associated absorbers of
background quasars will not prefer the systemic velocity of the foreground quasars, and hence will not affect the
average absorption measured for the foreground quasars. The median pair velocity separation of the CII sample is
11140 km/s. Dividing the sample ino two bins of pair velocity separation does not result in meaningful comparison, due
to the smaller number of pairs in each bin. Hwever we see no evidence for a difference in the centroid's redshift of
the two bins.

>B) To demonstrate that the quasar variability effect can plausibly produce the observed shift from Hubble flow alone
>the authors need to calculate the distance to which the quasars can substantially contribute to the UV background and
>convert this distance to a line-of-sight velocity assuming Hubble flow. Most of the work for this has already been done
>for the calculation of g_UV for Figure 1 and Table 1 so only a few additional steps are needed. Implementing this
>calculation in the Monte Carlo approach in Section 3.3 would be particularly convincing.

The strength of absorption is not only affected by the strength of the ionizing radiation field which g_UV measures,
but also by the impact parameter of the pairs. As QPQ8 demonstrated, the strength of the absorption decreases with
impact parameter. Further, g_UV is calculated under the assumption that the quasar shines isotropically and
non-intermittently. It is therefore not straightforward to isolate the quasar variability effect by examining
any possible correlation between the strength of absorption and g_UV. We added more text in the revised manuscript to
explain how g_UV is calculated.

>C) To clarify the discussion the authors should:
>1) Rephrase or expand the sentence/paragraph: "As a test of reasonableness, we calculate that, for a quasar opening
>angle of 120...". It is not clear what calculation was performed and how it represents a test of reasonableness.
>2) Remove the sentence "Light from the background quasar will travel smaller distance to reach the gas behind the
>foreground quasar than the gas in front of it." It is repetitive and the subsequent sentence explains this physical
>picture more clearly.
>3) The last sentence of the paper is misleading because the observed velocity shift cannot be explained by
>gravitational motion.

We expanded on the test of an arbitrary, reasonable quasar opening angle. In this test, we removed all absorbers that
fall within the quasar opening angle in our Monte Carlo simulations. We found that for an opening angle of 120 degrees,
the observed centroid may be reproduced while the intrinsic dispersion is still wthin the modeling error.
We removed the sentence "Light from the background quasar will travel..." as the referee suggested. This indeed helps
clear up the test. We rewrote the last sentence to separately provide the explanation for the large velocity widths
and the explanation for the average shift.

>Finally, I would like to thank the authors for going the extra mile and researching color choices that improve figure
>legibility for color blind readers. However, the stated change does not appear to have been implemented in Figure 6.
>Since the work has already been done, implementing the color change would be an easy correction.

We examined the cartoon figures again. A lighter shade of blue had indeed been used for the CIV ions, and it seems the
poor legibility comes from the figures being too compact. We enlarged the figures in the revised manuscript. We are
very happy to use another color scheme should the referee or editor have better suggestion.

### 2018-01-26 QPQ9/Submit/qpq9_resubmit.tex
small changes

In the following, we explore two possible explanations for the putative non-dynamical process that
provides the asymmetry.

One possibility is an asymmetric radiation field that preferentially ionizes the gas moving toward 
the observer, where the quasar is known to shine. Alternatively, the asymmetric radiation field 
may preferentially ionizes the gas at smaller Hubble velocity than the quasar. In 
Figure~\ref{fig:monopolar}, we show a cartoon of a quasar that is blocked in the direction 
pointing away from the observer. The gas observed in absorption preferentially lies behind the 
quasar. The asymmetric absorption arises from a transverse proxmity effect. 
\cite{Roos+15} and \cite{GaborBournaud14} performed simulations of a high-redshift disk galaxy 
including thermal AGN feedback and calculated radiative transfer in post-processing. They found 
the ionization radiation is typically asymmetric, due to either a dense clump that lies on one 
edge of the black hole or the black hole's location being slightly above the disk.
{\bf \cite{FaucherGiguere+16} represents the only simulation so far that is able to reproduce the
substantial amount of cool gas in quasar-mass haloes. We eagerly await their group to compare the
fraction of gas in inflows and outflows.}

@@

Another possible explanation arises from the finite lifetime of quasar episodes. 
Figure~\ref{fig:lighttravel} presents a cartoon for a light travel time argument. Light from the 
background quasar will travel smaller distance to reach the gas behind the foreground quasar 
than the gas in front of it. The light from the background quasar may arrive at the gas behind 
the foreground quasar before the ionizing radiation from the foreground quasar arrives. In 
QPQ7 and QPQ8, we reported that the enhancement in metal ion absorption relative to IGM average 
extends to at least $\approx200$\,kpc. This distance corresponds to a light travelling time of 
$\sim10^6$\,yr, which lies within existing constraints of quasar lifetime from observations 
\citep[e.g.][]{Martini04} and simulations \citep[e.g.][]{Hopkins+05}.

>>>>

In the following, we explore two possible explanations for the putative non-dynamical process that
provides the asymmetry. 
{\bf The explanations arise from a transverse proximity effect (QPQ4), which is the suppression of 
opacity in background sightlines passing close to foreground quasars.}

One possibility is an asymmetric radiation field that preferentially ionizes the gas moving toward 
the observer, where the quasar is known to shine. Alternatively, the asymmetric radiation field 
may preferentially ionizes the gas at smaller Hubble velocity than the quasar. In 
Figure~\ref{fig:monopolar}, we show a cartoon of a quasar that is blocked in the direction 
pointing away from the observer. The gas observed in absorption preferentially lies behind the 
quasar. 
\cite{Roos+15} and \cite{GaborBournaud14} performed simulations of a high-redshift disk galaxy 
including thermal AGN feedback and calculated radiative transfer in post-processing. They found 
the ionization radiation is typically asymmetric, due to either a dense clump that lies on one 
edge of the black hole or the black hole's location being slightly above the disk.
{\bf \cite{FaucherGiguere+16} represents the only simulation so far that is able to reproduce the
substantial amount of cool gas in quasar-mass haloes. We eagerly await their group to compare the
fraction of gas in inflows and outflows.}

@@

Another possible explanation arises from the finite lifetime of quasar episodes. 
Figure~\ref{fig:lighttravel} presents a cartoon for a light travel time argument. Light from the 
background quasar will travel smaller distance to reach the gas behind the foreground quasar 
than the gas in front of it. The light from the background quasar may arrive at the gas behind 
the foreground quasar before the ionizing radiation from the foreground quasar arrives. In 
QPQ7 and QPQ8, we reported that the enhancement in metal ion absorption relative to IGM average 
extends to at least $\approx200$\,kpc. This distance corresponds to a light travelling time of 
$\sim10^6$\,yr.
{\bf In turn, this implies a minimum quasar lifetime of $10^6$\,yr, which lies within existing
constraints from}
observations \citep[e.g.][]{Martini04} and simulations \citep[e.g.][]{Hopkins+05}.
